Model-based Clustering for Neural Populations

### 2021-10-04 documents/cluster_neural_population.tex
resolve clustering issues

@@

\section{TODO}
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
%	Think about MV-(Poisson)-GLM. Check \href{https://www.tandfonline.com/doi/full/10.1080/03610926.2012.743565?journalCode=lsta20}{ref1} and \href{https://www.tandfonline.com/doi/full/10.1080/02664763.2021.1877637?src=recsys}{ref2}.
	Improve the NUTS.
	\item
	Debug the clustering.
	\item
	Check and clear typo.
%	\item
%	After resolving (1), see if the MCMC is still trapped in local optima sometimes. Now, the chain seems will get stuck in local mode. Maybe updating the loading as a whole will remedy the problem a bit.
	\item
	Implement the mixture of finite mixture (MFM) by Jeff Miller.
	\item
	Find a more efficient way to generate new cluster parameters, otherwise the newly generated ones will always be rejected.
%	\item
%	improve DPMM. In current brute force implementation, the number of potential clusters can even go beyond number of neurons (\(N\)). There are several improvements, e.g. \href{https://link.springer.com/article/10.1007/s11222-009-9150-y}{Kalli et al., 2011}, \href{http://proceedings.mlr.press/v37/gea15.html}{Ge et al., 2015} and \href{https://link.springer.com/article/10.1007/s11222-014-9471-3}{Hastie et al, 2015}. Check them later.
	\item
	After all of them are resolved, switch to GP version

>>>>

\section{Some Issues in clustering}
The first problem is how to update the unobserved components. In DP, this is the same as how to generate new components.

Traditionally, the unobserved ones are just sampled from prior. However, this doesn't work in my case, since the dimension of $\bm{X}$ is super high ($6\times 1000$) in simulation. It's nearly impossible to accept the newly generated components.

Here's my solution: other parameters are generated from prior, but $\bm{X}$ in new component is just the same  as, or just another sample for the largest cluster.

In DP, we can do things even more aggressive: at the first few steps (e.g. pre-burn-in = 10), let all newly generated component parameters be the same (or samples from the same distribution) as the largest cluster. \textbf{This strategy encourages more clusters but keeps the clustering structure at the same time}.

After pre-burn-in, we goes back to normal: all parameters except for $\bm{X}$ are generated from priors. This gives the algorithm some chances to hit a "good" one, but the probability is not as high as in pre-burn-in.

The second problem comes when I let $\bm{Q}$ be full. For example, at one step, I have cluster index 1, 2, 4. But the sub-matrix of $\bm{Q}$ buying 1, 2, 4 columns and rows is not necessarily positive-definite. Here, I just do the easy thing: if the sub-matrix is not p.d., only keep the diagonal elements or block diagonal elements.

@@

\section{TODO}
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	Improve the NUTS.
	\item
	Debug the clustering.
	\item
	Check and clear typo.
%	\item
%	After resolving (1), see if the MCMC is still trapped in local optima sometimes. Now, the chain seems will get stuck in local mode. Maybe updating the loading as a whole will remedy the problem a bit.
	\item
	Implement the mixture of finite mixture (MFM) by Jeff Miller.
	\item
	Find a more efficient way to generate new cluster parameters, otherwise the newly generated ones will always be rejected. (\textbf{resolved}?)
%	\item
%	improve DPMM. In current brute force implementation, the number of potential clusters can even go beyond number of neurons (\(N\)). There are several improvements, e.g. \href{https://link.springer.com/article/10.1007/s11222-009-9150-y}{Kalli et al., 2011}, \href{http://proceedings.mlr.press/v37/gea15.html}{Ge et al., 2015} and \href{https://link.springer.com/article/10.1007/s11222-014-9471-3}{Hastie et al, 2015}. Check them later.
	\item
	After all of them are resolved, switch to GP version
